Quantum Mechanics Assignment Five

### 2016-02-12 quantum/latex/a5.tex
Add 15.tex
errthang

@@

>>>>

\section{1: Effect of E-field on H spectrum}
Here we examine how a perturbing electric field influences the n=2 to n=1 transition for
the Hydrogen atom.\\

\textbf{a.)} First we find the unperturbed transition energy:\\

\begin{align*}
  E_2 - E_1 &= -13.6 \mbox{eV} \left(\frac{1}{2^2} - \frac{1}{1^2}\right) = 10.2 \mbox{eV}\\
\end{align*}

\textbf{b.)} Now we apply a 0.5 V/m electric field and see how the spectrum changes. The Stark
Effect tells us that the energy levels will be perturbed by either zero or
$\pm3e\epsilon a_0$, which for this electric field would be $\pm 2.64*10^{-2}eV$. This comes out
to be a change of roughly $0.25$ percent, which is a fairly small value.\\

The perturbation due to the electric field can be interpreted as the energy change of the
electric dipole moment interacting with the applied field. Some of the n=2 states are radially
symmetric, and so have no electric dipole to alter the energy. However certain combinations of
orbitals (ie $\bra{200}+\bra{210}$) do have a bias to part of the electron, and so do have an
energy change.\\

@@

\section{11.9: Commutators}
Here we show that the spin magnitude operators \textbf{$S^2$} and \textbf{$I^2$} commute with
the perturbation Hamiltonian, while the component operators \textbf{$S_z$} and \textbf{$I_z$}
do not.

\begin{align*}
  \left[S^2, H'_{hf}\right] &= \left(S^2H'_{hf} - H'_{hf}S^2\right)\ket{m_p,m_s}\\
  &= S^2H'_{hf}\ket{m_p,m_s} - H'_{hf}S^2\ket{m_p,m_s}\\
  &= S^2E_{m_p,m_s}\ket{m_p,m_s} - H'_{hf}S^2\frac{3\hbar^2}{4}\ket{m_p,m_s}\\
  &= \frac{3\hbar^2}{4}E_{m_p,m_s}\ket{m_p,m_s} - E_{m_p,m_s}S^2\frac{3\hbar^2}{4}\ket{m_p,m_s}\\
  &= 0\\
\end{align*}
